---
title: "Assignment 4"
author: "Marco Palombo"
date: "5/18/2022"
output: 
  html_document:
    code_folding: hide
---

```{r setup, include = FALSE, message = FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE)

library(tidyverse)
library(here)
library(janitor)
library(thematic)
library(scales)
library(equatiomatic)
```

```{r import data}
#import damages data set
damages_raw <- read_csv(here("damages.csv"))

# created new dataset and make some changes
damages <- damages_raw %>% 
#rename column damages because having it names the same thing as the dataset is super stupid
  mutate(harm = damages) %>% 
  select(-damages)

#import warming data set
warming_raw <- read_csv(here("warming.csv"))

#create new dataset
warming <- warming_raw

```

## Question 1
```{r question 1}
# first add a column to the dataset of damages squared because it seems lm wants that to do a quadratic fit
damages_fit <- damages %>% 
  mutate(warming2 = warming^2)

# perform fit
quad_fit <- lm(harm ~ warming + warming2 + 0, data = damages_fit)
summary(quad_fit)

quad_func <- function(warming){
  harm <- quad_fit$coefficients[1]*warming + quad_fit$coefficients[2]*warming^2
  return(harm)
}

# populate column of harm from fit function
damages_fit$harm_fit = quad_func(damages_fit$warming)

ggplot(data = damages_fit, aes(x = warming, y = harm))+
  geom_point(aes(x = warming, y = harm))+
  geom_smooth(method = "lm", formula = y ~ x + I(x^2))+
  labs(x = "Warming (Degrees Celcius)", y = "Damages (figure out units)")
  
  
```

**Figure 1: Damages vs Warming. Points are from given dataset. Blue line is quadratic fit with zero intercept.**

## Question 2

```{r}
#populate columns of warming with damages for pulse, no pulse, difference between pulse and no pulse, difference in damage per ton co2
#pulse
warming$pulse_damage <- quad_func(warming$warming_pulse)
#nopulse
warming$nopulse_damage <- quad_func(warming$warming_baseline)
# difference
warming$dif <- warming$pulse_damage - warming$nopulse_damage

# unclear how to interpret the per ton CO2 comment for final column. Does this mean divide the whole column by 35? Or is there some sort of temporal component. Given it is referred to as a one time release today I will assume the 35 is in the air at the beginning of the dataset and will therefore simply divide the difference by 35
warming$dif_per_ton <- warming$dif/35e9

# now generate the plots
# no pulse
ggplot(data = warming)+
  geom_point(aes(x = year, y = nopulse_damage))+
  labs(title = "no pulse", x = "Calendar year", y = "Damages (unknown units)", 
       caption = "Figure 2: damages from no pusle with respect to year")

# pulse
ggplot(data = warming)+
  geom_point(aes(x = year, y = pulse_damage))+
  labs(title = "pulse", x = "Calendar year", y = "Damages (unknown units)", 
       caption = "Figure 3: damages from pulse with respect to year")

# difference
ggplot(data = warming)+
  geom_point(aes(x = year, y = dif))+
  labs(title = "difference", x = "Calendar year", y = "Damages (unknown units)", 
       caption = "Figure 4: difference in damages with respect to year")

# difference per ton
ggplot(data = warming)+
  geom_point(aes(x = year, y = dif_per_ton))+
  labs(title = "difference per ton", x = "Calendar year", y = "Damages (unknown units)", 
       caption = "Figure 5: difference in damages per tone CO2 with respect to year")
```

## Question 3

```{r}
# first calculate the net damages up to 2100 due to single tone of CO2. this will be accomplished by summing the column defined as the difference per ton
net_harm_per_ton <- sum(warming$dif_per_ton)

#create function of net present value 
#inputs
#rate: discount rate (%)
#damage: total damages at time t
#time: time at which damages have accrued
#present_time: time at which net present value is calculated

discount_damages <- function(rate, damage, time, present_time){
  NPV <- damage/ (1 + (rate/100))^(time-present_time)
  return(NPV)
}

# initialize dataframe for discounting
discount <- data_frame(rates = seq(0,10, 0.1), SCC = seq(0,10, 0.1))

#use for loop to calculate discounted damages up to 2100 for discount rates from 0 to 10
for(x in 1:length(discount$rates)){
  #apply function from above and sum all elements in column to find net discounted cost
  discount$SCC[x] <- sum(discount_damages(discount$rates[x],warming$dif_per_ton, warming$year, 2022))
}

ggplot(data = discount)+
  geom_point(aes(x = rates, y = SCC))

```

**Figure 6: net present value of future damages between 2021 and 2100 at time 2021 discounted with respect to various discount rates**


## Question 4

```{r}
#determine r from the equation given
#define values
rho <- .001 
eta <- 2 
g <- .01

#plug into the Gordon Ramsay Rule
r_gordon_ramsey <- (rho + eta*g)*100

# now plug into the previous equation and sum to find SCC
discount_gordon_ramsey <- sum(discount_damages(r_gordon_ramsey,warming$dif_per_ton, warming$year, 2022))

#plot this dumb thing
ggplot(data = discount)+
  geom_point(aes(x = rates, y = SCC))+
  geom_point(x = r_gordon_ramsey, y = discount_gordon_ramsey, color = "blue", size = 4)
```

**Figure 7: Net present value of future damages between 2021 and 2100 at time 2021 discounted with respect to various discount rates. Blue point shows discount rate and SCC associated with recommendation made by The National Academies of Sciences, Engineering, and Medicine**

r value using this equation will be: `r r_gordon_ramsey` %

## Question 5
```{r}
# define chosen discount rate
discount_rate <- 2

# build up matrix
warming_policy <- warming_raw %>% 
  mutate(warming_a_1 = warming_baseline,
         warming_a_2 = 1.5*warming_baseline,
         warming_b = warming_baseline)

#enforce 1.29 degree limit on policy b warming
for(x in 1:length(warming_policy$warming_b)){
  if(warming_policy$warming_b[x]>1.29){
    warming_policy$warming_b[x] = 1.29
  }
}

# use quadratic fit function from above to damages from Policy A, for now keep both potential outcomes separated
warming_policy$dam_a_1 <- quad_func(warming_policy$warming_a_1)
warming_policy$dam_a_2 <- quad_func(warming_policy$warming_a_2)

# do the same for policy B
warming_policy$dam_b <- quad_func(warming_policy$warming_b)


# apply discount rate and sum for years 2022-2100
# policy A option 1
discount_damages_a_1 <- sum(discount_damages(discount_rate,warming_policy$dam_a_1, warming$year, 2022))
# policy a option 2
discount_damages_a_2 <- sum(discount_damages(discount_rate,warming_policy$dam_a_2, warming$year, 2022))
# policy b
discount_damages_b <- sum(discount_damages(discount_rate,warming_policy$dam_b, warming$year, 2022))

# account for probabilities of two potential policy A outcomes
pol_a_1_prob <- .5
pol_a_2_prob <- .5

discount_damages_a <- pol_a_1_prob*discount_damages_a_1 + pol_a_2_prob*discount_damages_a_2

#create and populate dataframe incorporating the costs of policy implementation
damages_plot <- data_frame(implementation_cost = seq(0,5e15, by = 1e12))
damages_plot$pol_a <- discount_damages_a*rep(1,length(damages_plot$implementation_cost))
damages_plot$pol_b <- discount_damages_b*rep(1,length(damages_plot$implementation_cost))+damages_plot$implementation_cost
damages_plot$b_cross<- discount_damages_b*rep(1,length(damages_plot$implementation_cost))
  
  

# make this stupid figure
ggplot(data = damages_plot)+
geom_line(aes(x = implementation_cost, y = pol_a))+
  geom_line(aes(x = implementation_cost, y = pol_b))

root_func <- function(x){discount_damages_a-(discount_damages_b+x)}
intersection <- uniroot(root_func,c(2.0e15,2.5e15))$root
```

Present value damages policy A: `r discount_damages_a` units
Present value damages policy B: `r discount_damages_b` units

Congress should use Policy B if it costs less than $`r intersection`. However is policy B's cost exceeds this threshold policy A is preferred.

## Question 6

